---
title: "Examining Covariations Between Variables"
author: "Pauline I. Alvarado"
date: "Spring 2018"
output: pdf_document
---

This coding exercise was from the Program Evaluation and Data Analysis course taught by Dr. Nelson Lim at the University of Pennsylvania. Data was provided by the instructor.

# Key Packages
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  library(tidyverse)
  library(readr)
  library(ggplot2)
  library(ggthemes) 

```


# Bi-Variate or Multivariate Covariations
First step toward determine causality. Important to determine input-output, independent-dependent relationship between the variables, and design the visualization and analysis based on that determinant.

# Covariation between continuous variables

## Create scatterplot
Visualize covariation between the representation of females in a college major and the median income of the the major. Data taken from fivethirtyeight.
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}
  
  college_major <- read.csv("recent-grads.csv")

  ggplot(data = college_major) + geom_point(mapping = aes(x=ShareWomen, y=Median), color = "green") +
                                
        # Format x and y labels
          scale_x_continuous(name = "Percent Female", labels = scales::percent) +
          scale_y_continuous(name = "Median Wage", labels = scales::dollar)
    
```


## Add third dimension to the plot
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  ggplot(college_major) + geom_point(aes(x = ShareWomen, y = Median, color = Major_category))

```


## Facets
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}
  
  ggplot(data = college_major) +
      geom_point(mapping = aes (x = ShareWomen, y = Median, color = Major_category)) +
      facet_wrap(~Major_category, nrow = 2)

```



## Multivariate scatterplot with regression lines
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

 ggplot(data = college_major) +
    geom_point(mapping = aes(x=ShareWomen, y=Median, color=Major_category)) +
    geom_smooth(aes(x=ShareWomen, y=Median))

```



## Boxplot
Bin continuous variable to act like a categorical variable. Plot covariation between degree and wage. Data taken from the US Bureau of Labor Statistics' Current Population Survey.
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  cps_small <- readRDS("cps-2017-small.rds")

  # Adjust bin for "Yrs_Schooling"
    cps_small <- cps_small %>% drop_na(Yrs_Schooling) %>%
                               mutate(Degree = cut(Yrs_Schooling, breaks=c(-Inf, 11, 12, 15, 16, 20), 
                                      labels = c("<HS","HS","Some College", "BS", "Grad")))
  
  # Plot
    ggplot(data = cps_small, mapping = aes(x = Degree, y = Wage)) + 
      geom_boxplot() +
      scale_x_discrete(name = "Education") + 
      scale_y_continuous(name = "Wage", labels = scales::dollar)
    
```



# Remove the outliers and zoom in and see the relationship better
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  ggplot(data = cps_small, mapping = aes(x = Degree, y = Wage)) + 
    geom_boxplot(outlier.shape = NA) +
    scale_x_discrete(name = "Education") + 
    scale_y_continuous(name = "Wage", labels = scales::dollar, limits = c(0, 200000))

```



# Covariation Between Categorical & Continuous Variable
Explore the distribution of a continuous variable broken down by categorical variable.

## Density plot
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  #categorical = input/x
  #continuous = output/y
  
  ggplot(mpg, aes(hwy, fill=class, colour=class)) + geom_density(alpha=0.1)

```



## Box plot
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  ggplot(data = college_major) +
     
    # sort majors highest to lowest median salary
    geom_boxplot(mapping = aes(x=reorder(Major_category, Median, FUN=median), y=Median)) +
    
    #flip the axes
    coord_flip() + #flip the axes

    #remove scientific notation
    scale_y_continuous(labels=scales::dollar)

```



# Covariation between two categorical variables

## Circles
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  ggplot(data = mpg) + geom_count(mapping = aes(x=class, y=drv))

```



## Tiles
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  mpg %>%
    count(class,drv) %>%
    ggplot(mapping=aes(x=class, y=drv)) +
    geom_tile(mapping = aes(fill=n))

```
 
 
 
## Stacked Bar Graph
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  ggplot(data = mpg) + geom_bar(mapping = aes(x=class, fill=trans))

```
    


# Formatting Exercise
Difference in wage distributions across states
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  ggplot(data = cps_small) +
      geom_smooth(mapping = aes(x = Yrs_Schooling, y = Wage, colour = Sex)) +
    
    #add title, subtitle, caption
    labs(
      title = "Gender gap in wages increase with the level of education in 2017",
      subtitle = "On average, women earn less than men across all levels of education",
      caption = "Source: Current Population Survey") +
    
    #add annotations
    annotate("text", x = 8, y = 55000, label = "Non-linear regression lines with standard errors") +
    
    #format labels
    scale_x_continuous(name = "Education (Years of Schooling)", breaks = seq(0, 20, 2)) +
    scale_y_continuous(name = "Wage", breaks = seq(0, 150000, 25000), labels = scales::dollar) +
  
    #format line color
    scale_color_manual(values = c(Male="red", Female="blue"), labels=c("Men", "Women")) +
      
    #format grid lines
    theme(
      panel.grid.major = element_line(size=.5, linetype = 'solid', colour = "White"),
      panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "White")
    )


```





