---
title: "Visualizing Generalized Linear Models and Generalized Additive Models"
author: "Pauline I. Alvarado"
date: "Spring 2018"
output: pdf_document
---

This coding exercise was from the Program Evaluation and Data Analysis course taught by Dr. Nelson Lim at the University of Pennsylvania. Data was provided by the instructor.

# Data Management
## Load data and packages
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  library(tidyverse)
  library(ggplot2)
  library(readstata13)
  library(modelr)

  acs <- read.dta13("acsphillylaborforce.dta")

```



## Create duplicate viarables for better graphic & summary tables
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  acs <- acs %>% rename(Race = raceth,
                        Sex  = sex,
                        Education = educ_year,
                        Degree = education,
                        Occupation = gen_occ,
                        Industry = ind_cat5,
                        Income = incwage,
                        Managers = leader_cat,
                        Age = age,
                        College_major = major1,
                        Marital_status = marst) %>% 
                 mutate(Age = as.numeric(Age),
                        Age_sq = Age*Age,
                        Education11 = case_when((Education < 11) ~ 0,
                                                (Education >= 11) ~ Education - 11),
                        Education16 = case_when((Education < 16) ~ 0,
                                                (Education >= 16) ~ Education - 16))

```



# Generalized Linear Models
## Create linear models
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  # NOTE: family = gaussian (normal distribution)

  fit1 <- glm(Income ~ Sex, family = gaussian(link = identity), data=acs)

  
  fit2 <- glm(Income ~ Sex + Race, family = gaussian(link = identity), data = acs)
  
  fit3 <- glm(Income ~ Sex + Race + Education + Education11 + Education16, 
              family = gaussian(link = identity), data = acs)
  
  fit4 <- glm(Income ~ Sex + Race + Education + Education11 + Education16 +
              Age + Age_sq, family = gaussian(link = identity), data = acs)
  
  fit5 <- glm(Income ~ Sex + Race + Education + Education11 + Education16 +
              Age + Age_sq + Managers, family = gaussian(link = identity), data = acs)

```



## Compare models
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  anova(fit1, fit5, test = "Chisq")

```



## Visualize models in tables
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  library("texreg")
  screenreg(list(fit1,fit2,fit3))

```



## Use Packages to Automate Work: Spline Functions
Breaks down independent variables into a small number of segments connected by knots.
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  library(splines)

  fit6 <- glm(Income ~ Sex + Race + bs(Education,3) + poly(Age,2) + Managers, 
              family = gaussian(link = identity), data = acs)
  summary(fit6)


```



## Create plots
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  library(effects)

  plot(effect("bs(Education,3)", fit6))
  plot(effect("poly(Age,2)", fit6))
  plot(effect("Sex", fit6))
  plot(effect("Race", fit6))
  
  plot(allEffects(fit6, xlevels = 50))

```



# Generalized Additive Model
Only assume variables are additive and not linear. More flexible regression models.
## Create models and view summaries
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}
library(mgcv)
    fit_gam1 <- gam(Income ~ Sex + Race + Education + Age + Managers, data = acs)
    summary(fit_gam1) 
    
    fit_gam2 <- gam(Income ~ Sex + Race + s(Education) + s(Age) + Managers, data = acs)
    summary(fit_gam2)

```



## Plot
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  plot(fit_gam2, all.terms = TRUE)

```



# Logistic Regression
Estimate regression models predicting a categorical outcome
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  # Recode "Managers" to create a binary dependent variable
    acs <- acs  %>% mutate(manager = (as.numeric(fct_collapse(Managers,
                               yes = c("Chief Executives", "Managers", "Supervisors"),
                               no = "Other"))))

  acs$manager[acs$manager == 2] <- 0


```



## GLM
### View summary
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  fit7 <- glm(manager ~ Sex + Race + bs(Education,3) + bs(Age,3), 
            family = binomial(), data = acs)
  summary(fit7)

```

### Plot
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  plot(allEffects(fit7, xlevels = 50))
  plot(effect("bs(Education,3)", fit7))
  plot(effect("Sex", fit7))
  plot(effect("Race", fit7))

```



## GAM
### View summary
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  fit_gam3 <- gam(manager ~ Sex + Race + s(Education) + s(Age), data = acs)
  summary(fit_gam3)

```

### Plot
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  plot(fit_gam3, all.terms = TRUE)

```





































