---
title: "Power Analysis"
author: "Pauline I. Alvarado"
date: "Fall 2017"
output: pdf_document
---

This coding exercise was from the Statistics for Public Policy course taught by Dr. Matthew Levendusky at the University of Pennsylvania. Data was provided by the instructor.

# Definition
Probability to conclude that there is an effect to be found. In the real world, power analysis can help decide if the study is worth conducting. 1-power is the probability of a type II error (don't find an effect when there is an effect to found, e.g. pregnant lady getting a negative pregnancy test).

# Power.t.test() function
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}
  
  # power.t.test(n = NULL, 
               # delta = NULL, 
               # sd = 1, 
               # sig.level = 0.05,
               # power = NULL,
               # type = c("two.sample", "one.sample", "paired"),
               # alternative = c("two.sided", "one.sided"))

```



# Problem 1a
Average 3rd grader scores 65/100 on a reading test, with a standard deviation of 12 points. An after-school tutoring program claims to boost test socres by 5 points on average. If a pilot program of n = 10 students is conducted, what is the progability of a type-II error (alpha = 0.05 and H0: mu = 0 vs HA: mu > 0).
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  power.t.test(n = 10, 
               delta = 5, 
               sd = 12, 
               sig.level = 0.05,
               power = NULL,
               type = "one.sample",
               alternative = "one.sided")

  # probability of a type-II error
    1 - 0.33501

```



# Problem 1b
Increase sample size
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

 power.t.test(n = 50, 
               delta = 5, 
               sd = 12, 
               sig.level = 0.05,
               power = NULL,
               type = "one.sample",
               alternative = "one.sided")

  # probability of a type-II error
    1 - 0.8962296
    

```



# Problem 1c
In order to get the standard 0.80 power, how many students should be enrolled in the pilot?
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}

  power.t.test(n = NULL, 
               delta = 5, 
               sd = 12, 
               sig.level = 0.05,
               power = 0.80,
               type = "one.sample",
               alternative = "one.sided")


```



# Problem 2
Nonprofit program that mentors youths claims that youth spend less than 30 minutes a day doing risky activities. Evaluation study measures how many minutes per day the kids (n = 40) spend doing "risky activities." The average is 25 minutes per day on risky activities with a std deviation of 12 minutes.
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE, results='hold'}
  
  # Hypothesis:
    # H0: 30
    # HA: 25
    
  # What is the power?
    power.t.test(n = 40, 
               delta = 5, 
               sd = 12, 
               sig.level = 0.05,
               power = NULL,
               type = "one.sample",
               alternative = "two.sided")

    
  # Probability of a type II error.
    1 - 0.7290392
    
```

